\Large{Adversarial Training against Systematic Uncertainty}

### 2016-09-12 tex/paper.tex
save

    \node[black!20] at (1.5,-3) {...};

    % Adversary r
    \node at (9,0) {Adversary $r$};
    \draw (6.5,-0.5) rectangle (11.5,-5.5);

@@

    \node[black!20] at (9,-3) {...};

    \draw (14,-1.5) rectangle (17,-4.5);
    \path[->, shorten >= 0pt] (15.5,-0.5) edge (15.5,-1.5);
    \node[above] at (15.5,-0.5) {$z$};
    \path[->, shorten >= 0pt] (15.5,-4.5) edge (15.5,-5.5);
    \node[below] at (15.5,-5.5) {$p_{\theta_r}(z|f(X;\theta_f))$};
    \node at (15.5,-3) {Truncated ${\cal N}$};

@@

As for generative adversarial networks, we propose to
train $f$ and $r$ simultaneously, which we carry out by considering
the value function
\begin{equation}
    E(\theta_f, \theta_r) = {\cal L}_f(\theta_f) - {\cal L}_r(\theta_f, \theta_r)
\end{equation}
that we optimize by finding the saddle point $(\smash{\hat\theta_f}, \smash{\hat\theta_r})$ such that
\begin{align}
    \smash{\hat\theta_f} &= \arg \min_{\theta_f} E(\theta_f, \smash{\hat\theta_r}), \label{eqn:min_thetaf} \\
    \smash{\hat\theta_r} &= \arg \max_{\theta_r} E(\smash{\hat\theta_f}, \theta_r) \label{eqn:max_thetar}.
\end{align}
Without loss of generality, the adversarial training procedure to obtain
$(\smash{\hat\theta_f}, \smash{\hat\theta_r})$ is formally presented in
Algorithm~\ref{alg:adversarial-training} in the case of a binary classifier $f :
\mathbb{R}^p \mapsto [0,1]$ modeling $p(Y=1|X)$. For reasons further explained
in Section~\ref{sec:theory}, ${\cal L}_f$ and ${\cal L}_r$  are respectively set to the
expected value of the
negative log-likelihood of $Y|X$ under $f$ and of of $Z|f(X;\theta_f)$ under
$r$:
\begin{align}
    {\cal L}_f(\theta_f) &= \mathbb{E}_{Y|X} [ -\log p_{\theta_f} (Y|X) ], \\
    {\cal L}_r(\theta_f, \theta_r) &= \mathbb{E}_{Z|f(X;\theta_f)} [-\log p_{\theta_r} (Z|f(X;\theta_f))].
\end{align}
The optimization algorithm consists in using stochastic gradient descent
alternatively for solving Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar}.

@@

Interestingly, let us emphasize that these results hold using only the 1D output
of $f$ (in the case of binary classification) as input to the adversary. We
could similarly enforce an intermediate representation of the data to be
pivotal, e.g. as in \citep{ganin2014unsupervised}, but this is in fact not necessary.

>>>>

    \node[black!30] at (1.5,-3) {...};

    % Adversary r
    \node at (11.75,0) {Adversary $r$};
    \draw (6.5,-0.5) rectangle (11.5,-5.5);

@@

    \node[black!30] at (9,-3) {...};

    \draw (14,-1.5) rectangle (17,-4.5);
    \path[->, shorten >= 0pt] (15.5,-0.5) edge (15.5,-1.5);
    \node[above] at (15.5,-0.5) {$Z$};
    \path[->, shorten >= 0pt] (15.5,-4.5) edge (15.5,-5.5);
    \node[below] at (15.5,-5.5) {$p_{\theta_r}(Z|f(X;\theta_f))$};
    \node at (15.5,-3) {Truncated ${\cal N}$};

@@

As for generative adversarial networks, we propose to
train $f$ and $r$ simultaneously, which we carry out by considering
the value function
\begin{equation}
    E(\theta_f, \theta_r) = {\cal L}_f(\theta_f) - {\cal L}_r(\theta_f, \theta_r)
\end{equation}
that we optimize by finding the saddle point $(\smash{\hat\theta_f}, \smash{\hat\theta_r})$ such that
\begin{align}
    \smash{\hat\theta_f} &= \arg \min_{\theta_f} E(\theta_f, \smash{\hat\theta_r}), \label{eqn:min_thetaf} \\
    \smash{\hat\theta_r} &= \arg \max_{\theta_r} E(\smash{\hat\theta_f}, \theta_r) \label{eqn:max_thetar}.
\end{align}
Without loss of generality, the adversarial training procedure to obtain
$(\smash{\hat\theta_f}, \smash{\hat\theta_r})$ is formally presented in
Algorithm~\ref{alg:adversarial-training} in the case of a binary classifier $f :
\mathbb{R}^p \mapsto [0,1]$ modeling $p(Y=1|X)$. For reasons further explained
in Section~\ref{sec:theory}, ${\cal L}_f$ and ${\cal L}_r$  are respectively set to the
expected value of the
negative log-likelihood of $Y|X$ under $f$ and of $Z|f(X;\theta_f)$ under
$r$:
\begin{align}
    {\cal L}_f(\theta_f) &= \mathbb{E}_{Y|X} [ -\log p_{\theta_f} (Y|X) ], \\
    {\cal L}_r(\theta_f, \theta_r) &= \mathbb{E}_{Z|f(X;\theta_f)} [-\log p_{\theta_r} (Z|f(X;\theta_f))].
\end{align}
The optimization algorithm consists in using stochastic gradient descent
alternatively for solving Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar}.

@@

Interestingly, let us emphasize that these results hold using only the (1D) output
of $f$ (in the case of binary classification) as input to the adversary. We
could similarly enforce an intermediate representation of the data to be
pivotal, e.g. as in \citep{ganin2014unsupervised}, but this is in fact not
necessary. \glnote{This also suggests we could use for $r$ a simpler, maybe
closed form, representation than a neural network (e.g. a GMM?).}
